Wigglyhedra

### 2024-06-11 draft/wigglyhedra.tex
Added references and more writing.

We now describe the full subcategory of \emph{linear complexes} of \(\mathcal{C}_m\).
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\)
as bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[P^{\bullet} = \bigoplus_{g,h,i} M_i\langle g \rangle[h],\]
where \(g\) is the internal grading shift, and \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).
Set \(P^h\) to be the direct sum of all summands that have homological grading shift \(h\).
That is, in the notation above, we have
\[P^{h} = \bigoplus_{g,i} M_i\langle g \rangle[h].\]

@@

\end{definition}
Let \(\mathcal{A}_m\) be the smallest full subcategory of \(\mathcal{C}_m\) that is closed under isomorphism, which contains all the linear complexes.
For any integer \(N\), let \(\mathcal{C}_m^{\le N} \subset \mathcal{C}_m\) be the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, containing all complexes of level \(\leq N\).
For any integer \(N\), let \(\mathcal{C}_m^{\geq N} \subset \mathcal{C}_m\) be the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, containing all complexes of level \(\geq N\).
Observe that if \(P^{\cdot}\) has level \(\ell\), then \(P^{\cdot}[1]\) has level \(\ell + 1\), and that \(P^{\cdot}[-1]\) has level \(\ell - 1\).
\asilata{The directions of the shifts are a mess.}
\begin{proposition}\label{prop:standard-heart}
  The category \(\mathcal{A}_m\) is abelian.
  Furthermore, it is the heart of a bounded \(t\)-structure on \(\mathcal{C}_m\), which we call the \emph{standard \(t\)-structure} on \(\mathcal{C}_m\).
\end{proposition}
\asilata{Find a reference for this and possibly a cleaner statement.}

Also recall the following statements for future use.
\begin{defprop}\label{defprop:cohomology}
  Let \((\mathcal{C}^{\le 0}, \mathcal{C}^{\geq 0})\) be a bounded \(t\)-structure on a triangulated category \(\mathcal{C}\) with heart \(\mathcal{A} = \mathcal{C}^{\leq 0} \cap \mathcal{C}^{\geq 0}\).
  For each integer \(i\), we have the \(i\)th truncation functors
  \[\tau^{\leq i} \colon \mathcal{C} \to \mathcal{C}^{\leq i}\]
  and
  \[\tau^{\geq i} \colon \mathcal{C} \to \mathcal{C}^{\geq i},\]
  which are left and right adjoint respectively\asilata{Check left vs right.} to the inclusion functors \(\mathcal{C}^{\leq 0} \to \mathcal{C}\) and \(\mathcal{C}^{\geq 0} \to \mathcal{C}\).
  The \(i\)th \defn{cohomology functor} is defined to be the composition \(\tau^{\leq i} \circ \tau^{\geq i}\).
  It is equal to the composition \(\tau^{\geq i} \circ \tau^{\leq i}\), and maps an object of \(\mathcal{C}\) to an object of \(\mathcal{A}\).
\end{defprop}
In the presence of a bounded \(t\)-structure, every object of \(\mathcal{C}\) has a canonical filtration whose factors are shifts of its cohomology objects.
\begin{defprop}\label{defprop:cohomology-filtration}
  Let \(\mathcal{C}\) be any triangulated category, and \(\mathcal{A} \subset \mathcal{C}\) the heart of a bounded \(t\)-structure on \(\mathcal{C}\).
  Let \(X\) be a non-zero object of \(\mathcal{C}\), and let \(k\) and \(l\) be the maximum and minimum integers respectively for which \(H^k(X)\) and \(H^{l}(X)\) are non-zero.
  \asilata{Find references for cohomology and cohomology filtration.}
  Then \(X\) has a unique \defn{cohomology filtration}, which is the unique filtration of the form
  \[
    \begin{tikzcd}[column sep=0.1cm]
      0  \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k+1} \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} && \cdots \arrow[dashed]{ul} &&H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]  
\end{defprop}

>>>>

We now describe the \(t\)-structure induced by the full subcategory of \emph{linear complexes} of \(\mathcal{C}_m\); a variant of this construction is described in~\cite[Section 2.5]{lic.que:21}.
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\)
as bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[P^{\bullet} = \bigoplus_{g,h,i} M_i\langle g \rangle[h],\]
where \(g\) is the internal grading shift, and \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).
Set \(P^h\) to be the direct sum of all summands that have homological grading shift \(h\).
That is, in the notation above, we have
\[P^{h} = \bigoplus_{g,i} M_i\langle g \rangle[h].\]

@@

\end{definition}
For any integer \(N\), let \(\mathcal{C}_m^{\le N} \subset \mathcal{C}_m\) be the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, containing all complexes of level \(\geq -N\).
For any integer \(N\), let \(\mathcal{C}_m^{\geq N} \subset \mathcal{C}_m\) be the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, containing all complexes of level \(\leq -N\).

We refer the reader to e.g.~\cite[1.3]{bel.ber.del:82} for the definition of a \(t\)-structure on a triangulated category.
The following proposition is easily checked using the calculations in~\cref{rem:generating-morphisms} and standard techniques involving categories of complexes; we omit the proof.
\begin{proposition}
  The pair of subcategories \((\mathcal{C}_m^{\leq 0}, \mathcal{C}_m^{\geq 0})\) forms a bounded \(t\)-structure on \(\mathcal{C}_m\).
\end{proposition}
We call this \(t\)-structure the \defn{standard \(t\)-structure} on \(\mathcal{C}_m\).
Set \(\mathcal{A}_m = \mathcal{C}^{\leq 0}_m \cap \mathcal{C}_m^{\geq 0}\); this is the smallest full subcategory of \(\mathcal{C}_m\) closed under isomorphism, which contains all linear complexes.
It follows that \(\mathcal{A}_m\) is abelian.

Also record the following statements for future use; the first may be found in~\cite[1.3]{bel.ber.del:82}
\begin{defprop}\label{defprop:cohomology}
  Let \((\mathcal{C}^{\le 0}, \mathcal{C}^{\geq 0})\) be a \(t\)-structure on a triangulated category \(\mathcal{C}\) with heart \(\mathcal{A} = \mathcal{C}^{\leq 0} \cap \mathcal{C}^{\geq 0}\).
  For each integer \(i\), we have the \(i\)th truncation functors
  \[\tau^{\leq i} \colon \mathcal{C} \to \mathcal{C}^{\leq i} \quad\text{and}\quad \tau^{\geq i} \colon \mathcal{C} \to \mathcal{C}^{\geq i},\]
  which are right and left adjoint respectively to the inclusion functors \(\mathcal{C}^{\leq 0} \to \mathcal{C}\) and \(\mathcal{C}^{\geq 0} \to \mathcal{C}\).
  The \(i\)th \defn{cohomology functor} is defined to be the composition \(\tau^{\leq i} \circ \tau^{\geq i}\).
  It is canonically isomorphic to the composition \(\tau^{\geq i} \circ \tau^{\leq i}\), and maps an object of \(\mathcal{C}\) to an object of \(\mathcal{A}\).
\end{defprop}
The next statement, which is easy to check from the definitions, is mentioned in~\cite[Section 3.3]{bayer2011tour}.
\begin{defprop}\label{defprop:cohomology-filtration}
  Let \(\mathcal{C}\) be any triangulated category, and \(\mathcal{A} \subset \mathcal{C}\) the heart of a bounded \(t\)-structure on \(\mathcal{C}\).
  Let \(X\) be a non-zero object of \(\mathcal{C}\), and let \(k\) and \(l\) be the maximum and minimum integers respectively for which \(H^k(X)\) and \(H^{l}(X)\) are non-zero.
  Then \(X\) has a unique \defn{cohomology filtration}, which is the unique filtration of the form
  \[
    \begin{tikzcd}[column sep=0.1cm]
      0  \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k+1} \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} && \cdots \arrow[dashed]{ul} &&H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]  
\end{defprop}
